---
title: "Exam 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exam 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

  
```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center'
  )
library(operationsMgmt)
library(tidyverse)
library(nomnoml)
theme_set(theme_light())
```
  

```{r setup, eval = FALSE}
library(operationsMgmt)
library(tidyverse)
library(nomnoml)
```


# Module review: Day 1 Review 

## Homework 

**Homework is to be sent to:** patterson.homework@msutexas.edu

## Syllabus overview 

Course content - content map of the semester. 

Video lectures CC >> mgmt-sci-video 

- Software for using program to complete homework. 

BEP >> Break even point. 

Management Science Programs for the IBM Personal Computer 

## Topics covered

- Product - concepts, issues, techniques 
- Cost, Volume, Profit Analysis 
- Capacity planning, supply chain management
- Operations research tools 
- Linear programming 
- PERT 
- Break even analysis 
- Transportation and inventory controls 
- Lean manufacturing 
- Just in time production 
- Total Quality Management 
- Theory of constraints 



# Operations Management 

## Faster, Better, Cheaper 

People (businesses) need to be able to deliver products faster, better, 
or cheaper. 

- Developing a competitive advantage 

Price always matter >> cost competitive (price)

- Lower prices do not always mean lower quality. 

### Philosophies governing the operations world

#### Faster >> Just in time (JIT)

Dr. Taiichi Ohono - Founder JIT Vice president of production at Toyota. 

#### Better >> Total Quality Management(TQM)

Dr. Edwards Deming - Philosopher contributor to TQM movement.

#### Cheaper >> Cost competitive 

Dr. Eli Goldratt - Father of the Theory of Constraints. 
- PHD in physics. 
- Wrote a novel called The Goal 


## Operations 

The study of production systems that produce goods or provide services. 

- Applies to many different types of businesses (not just manufacturing).

Operations will apply to all different types of businesses - 

- Manufacturing 
- Health care 
- Banking 
- Government 
- Transportation 
- Consulting 
- Education 
- Retail/ Wholesale 
- All organizations 


### Manufacturing 

Goods producing systems typically classified as durable and non-durable goods. Ie: 
Hard goods such as large capital expenditures vs everyday purchases (soft goods). 

- Manufacturing 
- Durable 
- Non-durable 
- Construction 
- Mining 
- Agricultural 

### Service Sector 

Everything that is not included in the Manufacturing sector (knowledge work).

- Financial 
- Real estate 
- Insurance 
- Health care 
- Accounting / consulting 
- Education 
- Transportation 
- Other (non-goods producing)

## I/O Model: input output model

```{r, echo = FALSE, fig.align='center'}
nomnoml::nomnoml('
#spacing:20
[input] --> [process]
[process] --> [output]
', width = 400, height = 200)
```




```{r, echo = FALSE, fig.center = TRUE}

nomnoml::nomnoml("
#direction: right
#spacing:40
[input |
- Raw material] --> [process WIP]

[process WIP] --> [output |
- Final goods]

[process WIP | - In the process of being built
- Resources/ factors of production]

#direction:down
[Feed back loop] --> [input]
[Feed back loop] --> [process WIP]
[output] --> [Feed back loop]

[output] --> [consumer (end user)]

", width = 400, height = 500)
```



### Secondary inputs: factors of production/ resources 

Part of the work in process portion of the input, output model. 

- Land
- Labor 
- Capital - Equipment that is used to produce another good 
  - Hammer, computer, useful life greater than 1 year
  - Plant & equipment (PP&E)
  
- Money >> purchases productive resources (not considered a capital
good)


## Organization chart 

Cornerstone of a business or organization.

```{r, echo = FALSE}
nomnoml::nomnoml('
#direction: down
        [general manager] --> [operations | Supply side: \n people making products.]
        [general manager] --> [marketing | Demand side: \n Selling of product]
        [general manager] --> [finance | Capital budgeting: \n Allocation of resources]
        ', height = 400, width = 500)
```



## Management philosophy 

### Just in Time (JIT)

- Smaller warehouses - don't carry as much raw materials or inventory 
as they used too. 
- Leads to less waste 
- Smaller more frequent shipments is better/ more efficient
- Inventory >> Asset on the balance sheet 
- As long as the inventory has a high velocity (turning over) there 
are no issues. 
- Stale inventory is marked down - provide an incentive to the customer
- Goods will be sold for less that their costs in extreme cases 

### Functions of management 

How managers spend their time. 

- Planning - determining a future course of action 
- Short term - planning inventory/ production 
- Long term - capital expenditures (PPE) 

- Organizing - Factors of production - putting together the pieces of 
the puzzle, land labor and capital. 

- Staffing -  People and seats - making sure you have the right people 
in the right seats. 

- Directing - day to day planning/ instructions (more or less frequent).

- Controlling -  Comparing the actual results to the plan. 
- KPIs - performance 


### Productivty 

- Key issues for operations today: 
   
   `output / input = productivity `

Critical for the economy: how can we do more with the same given inputs/ resources.

Example: 

|     `(standard labor hours / amount of time worked) * 100 = efficiency`  

|     `(total output / total hours) = labor productivity`

- Technology: changed the world and how we do business. 
- Work methods: technology we use to do our job (tools and techniques). 
Ties back to technology. 
- Choice of equipment: comparing equipment - cost benefit analysis
- Management - JIT, TQM, TOC - new philosophies and principles that
prevail in an organization. 


### Key Issues 

- Productivity = output/ input 
- Quality: quality of the organization, workforce, and product 
  - Inversely proportional to output 

- Government/ economic regulations: regulate economic behavior in the marketplace 
(antitrust, etc.). Management focuses more behavioral regulations. 
- OSHA (1971) Occupational, safety, and health administration: worker 
health and safety. 
- EEOC (1965) Equal, employment, opportunity, commission: looks at 
discrimination in the work place. Makes it illegal to discriminate 
in the work place. 
- EPA (1970) Environmental protection agency: regulates business' 
impact on the environment. 
- (RRC) Railroad commission: oil and gas production in the state
of Texas. 
- Liability: make safe products for consumers (seat belts, cars, etc.), 
professional and service liabilities (doctors, accountants). 
- De-industrialization: manufacturing has collapsed in the last 
fifty years. Businesses will operate where it is most economic to 
operate. 
- Short-Term Focus - US companies more short term focused 
- Long Term Focus - Japanese more long term focused 
- Technology and Automation
- Globalization - free trade, open trade, and open borders. 

## Classification of Production Systems 

- Degree of standardization - products can be standard (mass produced)
- Comes out of inventory with no input from the end user. 
- Custom products - Built to order 
- Generally cost more and will require some time delay. 

Depending on the type of product being produced will dictate 
the type of operations: 

- Types of operations 
- Physical - Goods producing sector 
- Storage/ transportation - roads, railroads, air, water
- Exchange - retail and wholesale trade/ banking 
- Communication/ entertainment - television, music, movies

- Manufacturing vs. Service - Interaction between buyers and sellers 
  - Manufacturing: buyer and seller will have no interaction 
  - Service sector: interaction between buyer and seller (person to 
person relationship)


### Highly seasonable demand example

Smooth out the sales cycle by building inventory counter cyclically 
to peak demand. 

```{r}
tibble::tibble(
  qty = dnorm(seq(1, 100, 1), 40, 20) * 1000, 
  time = seq(as.Date('2015-02-01'), by = 1, length.out = 100),
) %>% 
  ggplot(aes(time, qty)) + 
  geom_line() + 
  labs(title = 'Graph of seasonal demand', 
       subtitle = 'Demand curve', 
       y = 'QTY', x = 'Time') + 
  theme_light() + 
  scale_x_date(date_labels =  '%b.\n%Y') + 
  geom_hline(yintercept = 14, color= 'red') + 
  geom_vline(xintercept = as.Date('2015-02-10') + 13, color= 'red') + 
  geom_vline(xintercept = as.Date('2015-02-10') + 47, color= 'red') + 
  geom_text(data = tibble(
    qty = 15, time = as.Date('2015-02-10')), 
    label = 'Factory\nCapacity', color = 'blue') + 
  geom_text(data = tibble(
    qty = c(2, 2, 2),
    time = seq(as.Date('2015-02-12'), by = 30, length.out = 3)), 
    label = c('Build inventory', 'Deplete Inventory', 'Build Inventory'))
```


Build products that fit the product cycle ie: (Flexible manufacturing)

- Build cooling equipment in the summer and re-tool (shift production) 
to heating equipment in colder months. 


Building inventory - Service sector 

- Knowledge workers sell their time and knowledge. 





# Break Even Analysis 

Analyzing revenue and costs 

TR: Total revenue 

- Income generated from normal business operations

TFC: Total fixed costs 

- Fixed costs remain the same regardless of business operations (rent)
- Costs that remain the same as output levels change 

TVC: Total variable costs 

- Costs that change depending on business operations (raw materials)
- Costs that fluctuate as output levels change

VC: Variable cost per unit 

- Contribution/ margin - variables costs to produce good or service. 
- Likely inversely proportional to output ie: as output increases the variable 
cost per unit will likely decrease if economies of scale are achieved. 

|     `TC = Total Fixed Costs + Total Variable Costs`

TRM: Total Raw Material 

- Variable cost - some input for creating/ manufacturing product

TDL: Total direct labor 

- Cost to manufacture a product or produce a good/ service 
- People directly involved to produce a product/ service 

NP: Net Profit <Loss> 

```
NP = TR - TC

Total Revenue = Volume (quantity) * Price 

Total Cost = Total Fixed + Total Variable Cost 

Net Profit = Total Revenue - Total Costs 
``` 


### Total variable costs: 

The cost to produce a good or service is fixed. 

- Raw materials 

```{r, fig.cap= 'Total Variable Costs'}
tibble(Quantity = 1:100, Cost = seq(100, 1000, length.out = 100)) %>% 
  ggplot(aes(Quantity, Cost)) + 
  geom_line() + 
  scale_y_continuous(labels = scales::dollar_format()) + 
  labs(title = "Total Variable Cost (TVC)", 
       subtitle = "Fixed cost per unit")
```


### Total fixed costs: 

Costs remain the same regardless of business activities, or output. 

- Rent 
- Service agreements 

```{r fig.cap= 'Total Fixed Costs'}
tibble(Quantity = 1:100, Cost = 4523) %>% 
  ggplot(aes(Quantity, Cost)) + 
  geom_line() + 
  scale_y_continuous(labels = scales::dollar_format()) + 
  labs(title = "Total Fixed Costs (TFC)"
  )
```

Relevant range: In reality, fixed costs are really only fixed within specific bounds 
or thresholds.   

Costs will not be absolutely fixed, but more generally can remain fixed up to 
a certain upper bound creating a stair step chart. 

- ie: Manufacturing output has increased so significantly, we need to expand into 
a larger factory. 

## Breaking down fixed and variable costs 

Has undergone changes in recent years. 

Fixed costs: 

- Rent 
- Salaries 
- Depreciation 
- Useful life of long term (capitalized) assets  
- Property tax 
- Most expenses 

Variable costs: 

- Raw material 
- Direct labor 
- Variable factory overhead
- Function of how much has been spent on direct labor  

Piece rate pay system - paid on the number of units produced. 

### New Philosophies 

TQM: Total quality management 
JIT: Just in time production 
TOC: Theory of constrains 

Fixed costs: 

- Salaries 
- Property tax 
- Depreciation 
- Rent 
- Direct labor 
- Most expenses 

Variable costs:

- Raw material: (maybe the only variable under new regime)

Maybe there are some holes in what is fixed vs. variable 

- Major difference being the definition of variable costs
- Is labor really a variable cost? 
- Production worker building products with a specific time constraint.  

Tenets of JIT - build only what you can sell! 

Regardless of whether you produce 40 or 50 units, you are still paid the same wage. 

```{r}
# 40 units / hour 
40 * (10 * 8)

# Next day 50 unit/ hour 
50 * (10 * 8)

# There is a difference in the cost per unit?
```


Next week - whether you are producing products, going through new training, or 
being told to stay home, you are still receiving the same wage. 

Generally, payroll does not fluctuate month to month regardless of business 
activity/ output. 

**Exam: Only true variable cost according to the new philosophies is raw material. Should consider payroll/ direct labor as a fixed cost.** 

Major takeaway is that companies will do things in different ways as long as they 
fit within accounting rules.

- Managerial accounting: internal decision making 
- Financial accounting: external reporting governed by financial accounting standards board
(FASB)

//TODO Throughput accounting? Activity based costing?

## Reviewing charts 

### ` Total cost = Variable Cost + Fixed costs`

If the price does not cover the cost of production, the producer/ supplier will 
lose money. The more you sell the more money you will lose. 

- Wheat framer (agriculture), oil & gas (lift cost/ variable cost)

```{r, fig.cap='Total costs'}
tibble(
  Quantity = 0:100, 
  TFC = 60, 
  TVC = Quantity * 1.6, 
  TC = TFC + TVC) %>% 
  tidyr::pivot_longer(cols = c('TFC', 'TVC', 'TC')) %>% 
  ggplot(aes(Quantity, value, color = name)) + 
  geom_line() + 
  labs(title = 'Total Costs', 
       x = 'Quantity', y = 'Cost') + 
  scale_y_continuous(label = scales::dollar_format()) + 
  geom_label(data = . %>% filter(Quantity == max(Quantity)),
             aes(label = name), position = 'identity') + 
  theme(legend.position = 'None')
```


### Break Even Point (BEP)

The point at which revenue will cover fixed costs and variable costs related to 
manufacturing a product. 

Might be true within a relevant range. As we sell products past the BEP, we will 
begin to produce positive cash flow. 

```{r, fig.cap='Price greater than variable costs'}
tibble(
  Quantity = 0:100, 
  price = 4,
  cost = 2.5, 
  TR = Quantity * price,
  TFC = 90, 
  TVC = Quantity * cost, 
  TC = TFC + TVC, 
  BEP_q = TFC / (price - cost),
  BEP_d = BEP_q * price
) %>% 
  tidyr::pivot_longer(cols = c('TFC', 'TVC', 'TC', 'TR')) %>% 
  ggplot(aes(Quantity, value, color = name)) + 
  geom_line() + 
  geom_label(
    data = . %>% filter(Quantity == max(Quantity)),
    aes(label = name), position = 'identity'
  ) + 
  geom_label(aes(mean(BEP_q)-15, mean(BEP_d) + 14, label = 'Break Even Point')) + 
  geom_segment(aes(x = 0, y = mean(BEP_d), yend = mean(BEP_d), xend = mean(BEP_q))) + 
  geom_segment(aes(x = mean(BEP_q), y = 0, yend = mean(BEP_d), xend = mean(BEP_q))) + 
  labs(title = 'Price > Variable Costs', 
       subtitle = 'Product or service produces positive cashflow', 
       x = 'Quantity', y = 'Cost') + 
  scale_y_continuous(label = scales::dollar_format()) + 
  theme(legend.position = 'None')
```


### Calculating net income 

```{r}
# Total fixed cost 
TFC = 6000 
# Variable costs 
VC = 5
# Price
P = 8 
output = 1500
# Total revenue 
TR = P * output # 12000 

# Total variable cost 
TVC = VC * output # 7500 
# Contribution margin
C = TR - TVC # 4500
# Total fixed costs 
TFC = 6000 

# Net loss 
NP = TR - TVC - TFC # 1500 

```


### BEP cont. 

Break even point is defined as: 
The number of units we have to sell to cover outlays:  

|     `BEP (units) = Total Fixed Costs / (Price - Variable Cost per Unit)`

Revenue generated to break even:  

|     `BEP (dollars) = BEP (units) * Price`


Margin of Safety: 

- The difference between total or expected sales, and the revenue that is 
required to break even.   

|     `Margin of Safety = Sales (Total or expected sales) - BEP (dollars)`

Contribution margin is defined as: 

|     `contribution margin ratio = (total revenue - variable costs) / total revenue` 
|     `contribution margin = Price per product - variable cost per product` 


Product markup: 

- Markup percentage of a product. 

|   `markup = (selling price - cost) / cost`



### Case study

Three variables to improve product profitability: 

- Fixed cost: attempt to reduce your fixed costs 
- Variable costs: look for other sources of raw materials 
- Price: raise the price 

#### Elasticity 

- Inelastic demand: not many substitutes, high switching costs, steep slope. 
- Elastic demand: many substitutes, price matters. 
- Unit elastic demand: change in price equally offset by the increase or decrease 
in the quantity demanded. 

```{r}
ggplot() + 
  geom_vline(xintercept = 200) + 
  labs()
```



# Break Even Analysis: Part 2 


## Example problems 


```
1. The Smith Manufacturing Company produces chairs. An analysis of their 
accounting data reveals:

Fixed cost $50,000 per year 
Variable cost $2 per chair
Capacity 20,000 chairs per year
Selling price $7 per chair
```

a. Compute the break-even point in number of chairs.


```{r}
tibble(
  FC = 50000, 
  VC = 2, 
  price = 7, 
  CAPACITY = 1:20000,
  TVC = CAPACITY * VC, 
  REV = CAPACITY * price, 
  TFC = TVC + FC) %>% 
  tidyr::pivot_longer(cols = c(TVC, REV, TFC)) %>%  
  ggplot(aes(CAPACITY, value, color = name)) + 
  geom_line() +
  geom_hline(aes(yintercept = FC)) +
  geom_point(aes(x = bep_units(VC, FC, price)[[1]], y = bep_dollars(VC, FC, price)[[1]])) + 
  scale_y_continuous(labels = scales::comma_format()) + 
  labs(title = "Calculating BEP") + 
  theme(legend.position = 'Bottom')


bep_units(2, 50000, 7)
bep_dollars(2, 50000, 7)

```


b. If sales (forecast or actual) = 12,000, what is the margin of safety?


```{r}
# Margin of Safety = Forecasted Sales - BEP Dollars 
scales::dollar(
  (12000 * 7) - bep_dollars(variable_cost = 2, fixed_cost = 50000, price = 7)
)
scales::comma(
  (12000) - bep_units(variable_cost = 2, fixed_cost = 50000, price = 7)
)
```

c.Find the number of chairs Smith must sell to show a profit of $30,000.

- Add the desired profit to the fixed cost

```{r}
# Number of units it takes to make 30,000 in profit 
bep_units(variable_cost = 2, fixed_cost =  50000 + 30000, price = 7)
```

d. What is the fixed cost per chair at 75 percent of capacity?

As output increases, fixed cost per unit will decrease as costs are spread over a 
larger output. 

```{r}
capacity <- 20000 
fixed_cost <- 50000 

# Fixed cost per chair 
fixed_cost / (capacity * .75)
```

Is the variable cost per unit constant? True 
Total variable cost will increase as output increases. 

Fixed costs per unit are inversely related to output. Fixed cost per unit will 
fall as output increase. 


```
2. The Taylor Feed Company produces feeds for chickens, hogs, cattle, and dogs. 
From available records we know the following:

Selling price                     Percent of $ sales
Feed for per ton    VC per ton    volume    
- Chickens    $30         $15         40%
- Hogs        $40         $15         20%
- Cattle      $36         $16         25%
- Dogs        $32         $12         15%

Annual fixed costs: $80,000
```

a . Find the total contribution per overall sales dollar with the
present product mix.

The current product mix offers a contribution margin % of 0.558. For every dollar 
sold, 0.56 is retained as gross profit. 

```{r}
problem2 <- tribble(
  ~price, ~cost, ~volume, 
  30, 15, .40,
  40, 15, .20,
  36, 16, .25,
  32, 12, .15
) %>% 
  mutate(
    margin = (price - cost) / price, 
    weighted_cm = margin * volume 
  ) 

problem2 %>% 
  summarise(cm_mix = sum(weighted_cm))
```


b. Find the break even in dollars

```{r}
# BEP Units = FC / (Price - VC )
# BEP Dollars = BEP Units * Price 
p2_bep <- problem2 %>% 
  summarise(cm = round(sum(weighted_cm), 2)) %>% 
  transmute(BEP_dollars = 80000 / cm) 

scales::dollar(p2_bep$BEP_dollars)
```


c. Find the margin of safety in $’s if $ Sales = $175,000 

```{r}
scales::dollar(175000 - p2_bep$BEP_dollars)
```


```
3. The Holmes Shoe Store currently stocks three lines of ladies'
shoes. Holmes is considering dropping one line of shoes and
adding two more. From the data given, decide whether or not
Holmes should make this change. Give your reasons.

Fixed costs: $30,000

SHOES NOW STOCKED: 
Type    Selling Price    VC per pair    Sales this year
- Flats   10                6             30,000
- Golf    16                12            10,000
- Dress   20                12            60,000


SHOES STOCKED IF PROPOSAL IS ACCEPTED: 
Type    Selling Price     VC per pair     Expected sales
- Flats     10                6                 25,000
- Dress     20                12                60,000
- Evening   16                8                 10,000
- Bedroom   6                 3                 5,000

Total $100,000

```



```{r}
current <- tribble(
  ~price, ~vc, ~sales, 
  10, 6, 30000,
  16, 12, 10000,
  20, 12, 60000) %>% 
  mutate(total = sum(sales),
         weight = sales / total)

proposal <- tribble(
  ~price, ~vc, ~sales, 
  10, 6, 25000, 
  20, 12, 60000, 
  16, 8, 10000, 
  6, 3, 5000) %>% 
  mutate(total = sum(sales),
         weight = sales / total)
```

The proposal offers a better product mix that will return a higher contribution 
margin. 

```{r}
# Calculating the average contribution margin for the product mix 
current %>% 
  mutate(cm = (price - vc) / price, 
         weighted_cm = weight * cm) %>% 
  summarise(average_cm = sum(weighted_cm))

proposal %>% 
  mutate(cm = (price - vc) / price, 
         weighted_cm = weight * cm) %>% 
  summarise(average_cm = sum(weighted_cm))
```


```
4. The Monopoly Supply Company is the sole supplier
of a patented line of widgets. Because their line of
widgets is sold to customers who require them as
replacement devices in expensive machinery, Monopoly
Supply experiences an inelastic demand pattern in sales
(each year they sell exactly 100,000 widgets regardless
of price). Recently, however, Monopoly Supply has been
threatened by anti trust actions and has learned, by the
grapevine, that they will be slapped with a lawsuit if
their profit on widgets next year exceeds 10 percent of
their fixed investment in widget production; this
investment is $5 million and is depreciated at $500,000
per year. If variable costs are $10 to make a widget, what
price should they ask next year in order to realize as
much profit as possible but still avoid antitrust litigation?
```

Perfectly inelastic meaning that the price of the product will not effect the 
quantity demanded (vertical line). 

An increase in price will not increase or decrease the number of products sold. 

```{r}
TFC=500000
# per widget
VC=10 

QTY=100000
INVESTMENT=5000000
MAXIMUM_PROFIT = TFC * .10 
```


```
5. A producer of electrical equipment is considering the
installation of one of two types of machines. A long run
sales forecast indicates that sales will not fall below
6,200 units per year for the next 5 years, the expected
life of each machine. Machine 1 will increase fixed costs
by $20,000 per year but will reduce variable costs by $6
per unit. Machine 2 will increase fixed costs by $4,000
per year but will reduce variable costs by $4 per unit.
Variable costs now amount to $20 per unit. At what
point are you indifferent as to which machine to
purchase? Which machine should be bought?
```

```{r}
t = 5 
forecast_units = 6200 
variable_costs = 20
```

Indifferent - what is the number of units at which the cost to operate each machine 
is equal to one another?

```{r}
total_cost <- function(fc, vc, units) fc + (vc * units)

# Machine 1: 
fc1 = 20000 
vc1 = variable_costs - 6
TC1 = t * total_cost(fc1, vc1, forecast_units)

# Machine 2: 
fc2 = 4000
vc2 = variable_costs - 4
TC2 = t * total_cost(fc2, vc2, forecast_units)

# Cost to operate is equal at 8,000 units 
(fc1 - fc2) / (vc2 - vc1)

```


```{r}
tibble(forecast_units = 0:10000) %>% 
  mutate(
    machine_1 = total_cost(fc1, vc1, forecast_units),
    machine_2 = total_cost(fc2, vc2, forecast_units)) %>% 
  tidyr::pivot_longer(cols = c(machine_1, machine_2)) %>% 
  ggplot(aes(forecast_units, value, color = name)) + 
  geom_line() + 
  labs(title = "Cost analysis", 
       subtitle = "Machine productivity", 
       x = 'Units produced', y = 'Total cost', 
       color = 'Machine:') + 
  scale_y_continuous(labels = scales::comma_format()) + 
  scale_x_continuous(labels = scales::comma_format()) + 
  theme(legend.position = 'bottom')
```


What is the cost of doing nothing?

Compare the cost of doing nothing ($0.00 fixed cost) with the machine that is closest
in comparison to fixed cost. This would be machine 2 since it has the lower 
fixed cost of $4,000.00. 

```{r}
tibble(forecast_units = 0:10000) %>% 
  mutate(
    current_machine = total_cost(0, 20, forecast_units),
    machine_1 = total_cost(fc1, vc1, forecast_units),
    machine_2 = total_cost(fc2, vc2, forecast_units)) %>% 
  tidyr::pivot_longer(cols = c(current_machine, machine_1, machine_2)) %>% 
  ggplot(aes(forecast_units, value, color = name)) + 
  geom_line() + 
  labs(title = "Cost analysis", 
       subtitle = "Machine productivity", 
       x = 'Units produced', y = 'Total cost', 
       color = 'Machine:') + 
  scale_y_continuous(labels = scales::comma_format()) + 
  scale_x_continuous(labels = scales::comma_format()) + 
  theme(legend.position = 'bottom')
```



```
8. Here are a firm's annual costs:
- Depreciation 40,000
- Salaries 53,000
- Materials used 30,000
- Advertising 15,000
- Direct labor 8,000
- Commissions on sales 16,000
- Taxes (property) 18,000

This firm sells six products; their total contribution per
overall sales dollar is 36 percent. What is the total
variable cost at the break even point?
```


```{r}
# Traditional orthodox view 
tibble(
  Depreciation = 40000,
  Salaries = 53000,
  Materials = 30000,
  Advertising = 15000,
  Direct = 8000,
  Commissions = 16000,
  Taxes = 18000) %>% gather() %>% 
  mutate(cost_type = c('F', 'F', 'V', 'F', 'V', 'V', 'F')) %>% 
  filter(cost_type == 'F') %>% 
  summarise(fixed_cost = sum(value)) %>% 
  # BEP Dollars 
  mutate(bep = fixed_cost / .36, 
         total_variable_cost = bep - fixed_cost) 
```


### Exam problem 

The only relevant cost in the short run is variable costs. 

```
9. The XYZ Railroad operates a train daily over an established route.
The railroad accountants have gathered the following cost figures
for different train lengths:________________________________

                         Total cost        Aver. cost per car
Engine and 10 cars            2,700                 270
Engine and 20 cars            3,200                 160
Engine and 30 cars            3,700                 123
Engine and 40 cars            4,200                 105
Engine and 50 cars (maximum)  4,700                 94

The operation is currently profitable at an average train length of 35
cars. The railroad competes with the Cannonball Truck Line over
the same route. The truck line approaches the railroad and offers to
pay it $86 per truck to haul trucks piggy back over the same route.
Each railroad flatcar will carry one truck, and the railroad has a
plentiful supply of flatcars. Extra costs would be $7.50 per truck to
cover the cost of loading and unloading. The trucking company will
not guarantee any minimum number of trucks to be hauled. Evaluate
this offer in terms of the profit opportunities. The train makes
300 round trips annually.
```

Should the rail company carry trucks at $86 per car?

What is the contribution margin per car without the trucking company and with 
the trucking company?

What is the fixed cost per car?
What is the variable cost per car?

```{r}
problem9 <- tribble(
  ~cars, ~tc, ~avg_cost_car,
  10, 2700, 270,
  20, 3200, 160,
  30, 3700, 123,
  40, 4200, 105,
  50, 4700, 94
) 

# Regression indicates 
# variable cost / car: 50 
# fixed cost: 2,200 
lm(tc ~ cars, data = problem9)
```

Start by calculating the variable cost per car. 

The change in the total cost from each bin is $500.00. (4,700 - 4,200 = 500)

```{r}
change_in_cars = 10 
change_in_price = 500 

# Variable cost per car 
cost_car = change_in_price / change_in_cars

```

The contribution margin is positive. 
| `86 - 57.50 = 28.50`




