---
title: "Homework Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{operationsMgmt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r, include = FALSE}
  knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
    )
```

Working homework assignments using R and python instead of expensive proprietary
closed source software. 

# Mgmt-Sci Video 

Working the examples from the mgmtsci video introducing the Management Science
Programs for the IBM Personal Computer. 

## Linear programming (Maximization) 

Solving the example problem in R using the {lpSolve} package.

Maximize 20B + 30C  
Constraints -- Subject to:   
B=Bearings C=Channels ~ Departments  
3B+2C<=16Machine  
1B+3C<=12Plate  
2B+2C<=6Polish  

- objective.in: Vector of coefficients of objective function  
- const.mat: One row per constraint; one column per variable  
- const.dir: Vector of character strings giving direction of the constraint  
- const.rhs: Vector of numeric constraints for RHS   

```{r}
library(lpSolve)

f.type <- 'max'
f.obj <- c(20, 30)
f.con <- matrix(c(3, 2, 1, 3, 2, 2), nrow = 3, byrow = TRUE)
f.dir <- c('<=', '<=', '<=')
f.rhs <- c(16, 12, 6)

lp(f.type, f.obj, f.con, f.dir, f.rhs)

```



## Project Evaluation Review Technique (PERT)

Examining the critical path. 
3 events (nodes)
1 time estimate per event (edges)

Diagramming the problem in R using {visNetwork}.

```{r}
library(visNetwork)
library(tibble)

list(
  nodes = tibble(
    id = 1:3, 
    label = paste('Node', 1:3), 
    shape = 'ellipse'
  ), 
  edges = tibble(
    from = c(1,2,1), to = c(2,3,3),
    label = c(8, 3, 14),
    font.color = c("red", "blue", "green"), 
    font.size = c(20,20, 20))
) %>% 
  {visNetwork(
    nodes = .[[1]],edges =  .[[2]], 
    main = 'Example 2 PERT: MgmtSciVideo', width = '100%')} %>% 
  visEdges(
    shadow = TRUE,
    arrows =list(to = list(enabled = TRUE, scaleFactor = 1)),
    color = list(color = "lightblue", highlight = "red"))

```


The above example has a critical path from 1 >> 3. 
The standard deviation cannot be calculate because of scalar time estimates.
Expected completion is 14. 

## Linear programming (Transportation Model) 

Solving the transportation model example problem using the {lpSolve} package with 
some assumptions. 

Unbalanced problem
3X5 Matrix 
Assuming capacity cannot be exceeded and demand must be satisfied. 

[Link](http://www.supplychaindataanalytics.com/solving-bronsons-transport-problem-with-lp-transport-in-r-using-lpsolve/) to helpful blog post discussing a possible solution.

```{r}

library(lpSolve)

# specifying cost matrix 
cost.mat <- matrix(nrow=3,ncol=5)
cost.mat[1,] <- c(5.6, 12, 14.5, 11, 0)
cost.mat[2,] <- c(11, 6.5, 4, 5, 0)
cost.mat[3,] <- c(100, 9, 5.5, 100, 0)

# this is a minimization problem
direction = "min"

# capacity may not be exceeded
row.signs <- rep("<=",3)
row.rhs <- c(7250,10150,4350)

# demand must be satisfied
col.signs <- rep(">=",5)
col.rhs <- c(8700,5800,2900,2175, 2175)

solution <- lp.transport(
  cost.mat = cost.mat,
  direction = direction,
  row.signs = row.signs,
  row.rhs = row.rhs,
  col.signs = col.signs,
  col.rhs = col.rhs
  )

cols_ <- c('NY', 'ATL', 'HOU', 'LA', 'DUMMY')
rows_ <- c('RC', 'SAL', 'WEST')

answer <- solution$solution 
rownames(answer) <- rows_
colnames(answer) <- cols_

answer %>% 
  data.frame() %>% 
  knitr::kable()
```

Shipping cost for the problem: 

```{r}
solution
```


Maximization of profit or minimization of costs. 


# Homework 1 (beppart2homework)

The homework assignment. 

