---
title: "Week 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Week 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

  
```{r, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```
  

```{r setup}
library(operationsMgmt)
library(tidyverse)
library(nomnoml)
```


# Module review 

## Homework 

**Homework is to be sent to:** patterson.homework@msutexas.edu

## Syllabus overview 

Course content - content map of the semester. 

Video lectures CC >> mgmt-sci-video 

- Software for using program to complete homework. 

BEP >> Break even point. 

Management Science Programs for the IBM Personal Computer 

## Topics covered

- Product - concepts, issues, techniques 
- Cost, Volume, Profit Analysis 
- Capacity planning, supply chain management
- Operations research tools 
- Linear programming 
- PERT 
- Break even analysis 
- Transportation and inventory controls 
- Lean manufacturing 
- Just in time production 
- Total Quality Management 
- Theory of constraints 



# Operations Management (Introdcution video)

## Faster, Better, Cheaper 

People (businesses) need to be able to deliver products faster, better, 
or cheaper. 

- Developing a competitive advantage 

Price always matter >> cost competitive (price)
- Lower prices do not always mean lower quality. 

### Philosophies governing the operations world

#### Faster >> Just in time (JIT)

Dr. Taiichi Ohono - Founder JIT Vice president of production at Toyota. 

#### Better >> Total Quality Management(TQM)

Dr. Edwards Deming - Philosopher contributor to TQM movement.

#### Cheaper >> Cost competitive 

Dr. Eli Goldratt - Father of the Theory of Constraints. 
- PHD in physics. 
- Wrote a novel called The Goal 


## Operations 

The study of production systems that product goods or provide services. 

- Applies to many different types of businesses (not just manufacturing).

Operations will apply to all different types of businesses - 

- Manufacturing 
- Health care 
- Banking 
- Government 
- Transportation 
- Consulting 
- Education 
- Retail/ Wholesale 
- All organizations 


### Manufacturing >> Goods producing systems 

- Manufacturing 
- Durable 
- Non-durable 
- Construction 
- Mining 
- Agricultural 

### Service Sector 

Everything that is not included in Manufacturing sector.

- Financial 
- Real estate 
- Insurance 
- Health care 
- Accounting / consulting 
- Education 
- Transportation 
- Other (non-goods producing)

## I/O Model: input output model

```{r, echo = FALSE}
nomnoml::nomnoml('
#spacing:40
[input] --> [process]
[process] --> [output]
')
```




```{r, echo = FALSE}

nomnoml::nomnoml("
#direction: right
#spacing:40
[input |
- Raw material] --> [process WIP]

[process WIP] --> [output |
- Final goods]

[process WIP | - In the process of being built
- Resources factors of production]

#direction:down
[Feed back loop] --> [input]
[Feed back loop] --> [process WIP]
[output] --> [Feed back loop]

[output] --> [consumer (end user)]

")
```



### Secondary inputs - factors of production / resources 

Part of the work in process portion of the input, output model. 

- Land
- Labor 
- Capital - good that is used to produce another good 
  - Hammer, computer, useful life greater than 1 year
  - Plant & equipment 
  - Money >> purchases productive resources (not considered a capital
good)


## Organization chart 

Cornerstone of a business or organization.

```{r, echo = FALSE}
nomnoml::nomnoml('
#direction: down
        [general manager] --> [operations | Supply side: \n people making products.]
        [general manager] --> [marketing | Demand side: \n Selling of product]
        [general manager] --> [finance | Capital budgeting: \n Allocation of resources]
        ')
```



## Management philosophy 

### Just in Time (JIT)


- Smaller warehouses - don't carry as much raw materials or inventory 
as they used too. 
- Leads to less waste 
- Smaller more frequent shipments is better/ more efficient
- Inventory >> Asset on the balance sheet 
- As long as the inventory has a high velocity (turning over) there 
are no issues. 
- Stale inventory is marked down - provide an incentive to the customer
- Goods will be sold for less that their costs in extreme cases 

### Functions of management 

How managers spend their time. 

- Planning - determining a future course of action 
- Short term - planning inventory/ production 
- Long term - capital expenditures (PPE) 

- Organizing - Factors of production - putting together the pieces of 
the puzzle, land labor and capital. 

- Staffing -  People and seats - making sure you have the right people 
in the right seats. 

- Directing - day to day planning/ instructions (more or less frequent).

- Controlling -  Comparing the actual results to the plan. 
- KPIs - performance 


### Productivty 

- Key issues for operations today: 
   
   `output / input = productivity `

Critical for the economy can we do more with the same give resources. 

- Technology - changed the world and how we do business. 
- Work methods -technology we use to do our job (tools and techniques). 
Ties back to technology. 
- Choice of equipment - comparing equipment - cost benefit analysis
- Management - JIT, TQM, TOC - new philosophies and principles that
prevail in an organization. 


### Key Issues 

- Productivity = output/ input 
- Quality - quality of the organization, workforce, and product 
  - Inversely proportional to output 

- Government Regulations - economic regulations - regulate economic 
behavior in the marketplace (antitrust, etc.). Management focuses more 
behavioral regulations. 
- OSHA (1971) Occupational, safety, and health administration - worker 
health and safety. 
- EEOC (1965) Equal, employment, opportunity, commission - looks at 
discrimination in the work place. Makes it illegal to discriminate 
in the work place. 
- EPA (1970) Environmental protection agency - regulates business' 
impact on the environment. 
- RRC - Railroad commission - oil and gas production in the state
of Texas. 

- Liability - make safe products for consumers (seat belts, cars, etc.), 
professional and service liabilities (doctors, accountants). 
- De-industrialization - manufacturing has collapsed in the last 
fifty years. Businesses will operate where it is most economic to 
operate. 

- Short-Term Focus - US companies more short term focused 
- Long Term Focus - Japanese more long term focused 
- Thinks long term is superior philosophy 

- Technology and Automation
- Globalization - free trade, open trade, and open borders. 

## Classification of Production Systems 

- Degree of standardization - products can be standard (mass produced)
- Comes out of inventory with no input from the end user. 
- Custom products - Built to order 
- Generally cost more and will require some time delay. 

Depending on the type of product being produced will dictate 
the type of operations: 

- Types of operations 
- Physical - Goods producing sector 
- Storage/ transportation - roads, railroads, air, water
- Exchange - retail and wholesale trade/ banking 
- Communication/ entertainment - television, music, movies

- Manufacturing vs. Service - Interaction between buyers and sellers 
  - Manufacturing: buyer and seller will have no interaction 
  - Service sector: interaction between buyer and seller (person to 
person relationship)


### Highly seasonable demand example

Smooth out the sales cycle by building inventory counter cyclically 
to peak demand. 

```{r, echo = FALSE}
tibble::tibble(
  qty = dnorm(seq(1, 100, 1), 40, 20) * 1000, 
  time = seq(as.Date('2015-02-01'), by = 1, length.out = 100),
) %>% 
  ggplot(aes(time, qty)) + 
  geom_line() + 
  labs(title = 'Graph of seasonal demand', 
       subtitle = 'Demand curve', 
       y = 'QTY', x = 'Time') + 
  theme_light() + 
  scale_x_date(date_labels =  '%b.\n%Y') + 
  geom_hline(yintercept = 14, color= 'red') + 
  geom_vline(xintercept = as.Date('2015-02-10') + 13, color= 'red') + 
  geom_vline(xintercept = as.Date('2015-02-10') + 47, color= 'red') + 
  geom_text(data = tibble(
    qty = 15, time = as.Date('2015-02-10')), 
    label = 'Factory\nCapacity', color = 'blue') + 
  geom_text(data = tibble(
    qty = c(2, 2, 2),
    time = seq(as.Date('2015-02-12'), by = 30, length.out = 3)), 
    label = c('Build inventory', 'Deplete Inventory', 'Build Inventory'))
```


Build products that fit the product cycle ie: (Flexible manufacturing)

- Build cooling equipment in the summer and re-tool (shift production) 
to heating equipment in colder months. 


Building inventory - Service sector 

- Knowledge workers sell their time and knowledge. 