---
title: "Homework 1"
author: "Bill Ash (5555WRA)"
date: "2/2/2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Operations Break-Even Examples using software

Formula for calculating break even point 

|   `BEP_Unit = FixedCost / (SellingPrice - VariableCost)` 
|   `ContributionMargin = (TotalRevenue - VariableCosts) / TotalRevenue` 
|   `BEP_Dollar = FixedCost / ContributionMargin`
|   `BEP_Dollar = BEP_Unit * SellingPrice` 

## BEP Functions

```{r}
# BEP Unit 
bep_unit <- function(fixed_cost, selling_price, variable_cost) {
  paste0('Break Even Point Units: ', 
         scales::comma(
           fixed_cost / (selling_price - variable_cost), accuracy = .01
         )
  )
}
# BEP Revenue
bep_rev <- function(fixed_cost, selling_price='', variable_cost='', cm = NULL) {
  
  if (is.null(cm)) { 
    cm <- ((selling_price - variable_cost) / selling_price)
  }
  
  paste0('Break Even Point: ', 
         scales::dollar(
           fixed_cost / cm , accuracy = .01
         )
  )
}
```



### Problem 1 

```
1. The Beta Manufacturing Company produces
stepladders that sell for $27 each. The production
capacity is limited to 48,000 units per year. Fixed
costs are $50,000 and variable costs are $12 per
item. Compute the break even point in number of
stepladders. How many items must be sold to
provide a profit of $20,000 What is allocation of
fixed costs per stepladder at 75 percent capacity?
```

```{r}
price = 27
capacity = 48000
variable_cost_unit = 12
total_fixed_cost = 50000 
expected_profit = 20000
```

### Answer 1    

Compute the break even point in number of
stepladders:   

>`r bep_unit(total_fixed_cost, price, variable_cost_unit)`

How many items must be sold to
provide a profit of $20,000:     

>`r bep_unit(sum(total_fixed_cost, expected_profit), price, variable_cost_unit)`

What is allocation of fixed costs per stepladder at 75 percent capacity?  

>`r scales::dollar(total_fixed_cost / (capacity * .75), accuracy=.01)` Fixed cost
per unit. 

### Problem 2   

```
2. A firm produces four models of tool sets. The fixed costs are
$42,000 and the other data are as follows: 

Model     Qty    Price  Variable Cost     Sales  Revenue
S     4,000     $20   6         $80,000
T     5,000     32    8        $160,000
U     1,000     60    20         $60,000
V     2,000     50    18         $100,000

- Total $400,000

Determine the break even in dollars

What is the Margin of Safety if Sales =$75,000?
```

```{r fig.cap='Table 2'}
prob_2 <- tibble::tribble(
  ~Model, ~Qty, ~Price, ~Variable_Cost, ~Sales_Revenue,
  'S', 4000, 20, 6, 80000,
  'T', 5000, 32, 8, 160000,
  'U', 1000, 60, 20, 60000,
  'V', 2000, 50, 18, 100000
)
fixed_costs = 42000
total_rev = 400000
knitr::kable(prob_2)
```

```{r}
cm_2 <- prob_2 %>% 
  mutate(
    cm = (Price - Variable_Cost) / Price, 
    relative_sales = Sales_Revenue / total_rev,
    total = cm * relative_sales
  ) %>% 
  summarise(cm = sum(total))
```

#### Answer 2:        



Determine the break even in dollars:   

> `r bep_rev(fixed_costs, cm = cm_2$cm)`  

What is the Margin of Safety if Sales = $75,000?  
  
>$15,000



### Problem 3   

```
3. The Clutch Engineering Company is proposing to locate
a branch office in one of two West Coast locations, A or B.
These two sites have quite different estimated operating
costs:
                            A        B
Engineering labor cost     $15/hr. $16/
Materials and supplies    2.40/hr. $1.80/
- (tied to engineering hours)
Variable overhead         5.50/hr. 4.40/hr.
Total annual fixed cost   $150,000 $190,000
Price to customers.       $30 $30

Consider the hourly costs to be variable costs.

a. Compute the break even for both locations.

b. At what level of output (number of hours) are you in
different to the location?
```

```{r}
prob_3 <- tibble::tibble(
  product = c('a', 'b'),
  v_labor_cost = c(15, 16), 
  v_materials = c(2.40,  1.80), 
  v_variable_overhead = c(5.50, 4.40), 
  total_fixed_cost = c(150000, 190000), 
  price = c(30,  30)
) 

knitr::kable(prob_3)
```

```{r}
resp <- prob_3 %>% 
  rowwise() %>% 
  mutate(variable_costs = sum(across(starts_with('v_'), sum)))
```


#### Answer 3      




>A: `r bep_unit(resp[1, 'total_fixed_cost'][[1]],resp[1, 'price'][[1]], resp[1, 'variable_costs'][[1]] )`

  
>B: `r bep_unit(resp[2, 'total_fixed_cost'][[1]],resp[2, 'price'][[1]], resp[2, 'variable_costs'][[1]] )`



Indifference:    

>`150,000 + 22.9x = 190,000 + 22.2x`    
`0.7x = 40,000`    
57,142 hours  

Bill Ash 