---
title: "MGMT-3453-X20: Homework 15-17"
author: "Bill Ash (5555WRA)"
date: "4/20/2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
library(tidyverse)
```



## Simulation (LONG PROBLEM)

Order cost == shipping: the extra cost associated with processing the order. 
- There is no carry cost if there is no inventory 
- If there is 0 units there is no stock out costs 
- If there is 0 units and we sell a unit there IS A stock out cost. 

```{r}
# Simulation problem: start with no open orders

# $1/ Unit / Week on the ending balance 
order_cost = 10 

# $3/ Unit/ Week 
carry_cost = 1

stock_out_cost = 3

# Current stock Units 
balance_on_hand = 12

# Reorder point 
order_point = 7

# EOQ
order_quantity = 8 
```


**Lead time analysis** - *the lead time is not constant*

```{r}
# Simulate the various lead times
lead_times <- tibble(
  lead_time = 1:5, 
  lead_frequency = c(2, 6, 11, 5, 1), 
  # probability the lead time will be 1 week or 5 weeks.
  lead_probability = lead_frequency / sum(lead_frequency), 
  lead_cdf = cumsum(lead_probability)
)

lead_times %>% 
  ggplot(aes(lead_time, lead_cdf)) + 
  geom_line() + 
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = '75% of all lead times were three weeks or less')
```


**Demand analysis**: *Demand is not always constant* 

- Have we ever had a situation in which we sold zero units: once 

```{r}
# Historical data 
demand_analysis <- tibble(
  # Units sold in a given week 
  demand_week = 0:6, 
  demand_frequency = c(1, 6, 9, 19, 8, 5, 2), 
  demand_probability = demand_frequency / sum(demand_frequency), 
  demand_cdf = cumsum(demand_probability)
)

demand_analysis %>% 
  ggplot(aes(demand_week, demand_cdf)) + 
  geom_line() + 
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(title = 'Probability three or more units\nsold in a given week is <= 70%', 
  x = 'Quantity demanded / week', y = 'Probability')
```


**Reorder point question**

How many do we have on hand / on order 

- `quantity on hand + quantity on order <= Reorder Point`
  - Yes, place an order 
  - No, don't place an order 


**SIMULATION EXPLANATION** 

```{r}
# Order point is 7: the point at which we need to order new product 
# Starting quantity on hand is 12.
# Week one we do not need to order. 

tibble(
  week = 1:10, 
  # Randomly chosen lead time 1:5 weeks 
  # LeadTimeRandom = hash function for choosing the lead time 
  # Randomly chosen demand 0:6 units / week 
  # DemandNumberRandom = hash function for choosing quantity demanded
  QuantityDemanded = c(1, 3, 2, 4, 2, 1, 3, 3, 0, 4),
  # Starting quantity of 12 - 1 = 11 for week one 
  QuantityOnHand = c(11, 8, 6, 2, 0, -1, 4, 1, 9, 5),
  # Order when TotalQuantity <= 7
  QuantityOrdered = c(0, 0, 0, 8, 0, 0, 8, 0, 0, 0), 
  
  # Lead time probability is used to determine how long the order is in transit
    # Could be a value 1:6. 
          OnOrder = c(0, 0, 0, 0, 8, 8, 0, 8, 0, 0), 
  # Product is received two weeks later 
  QuantityReceived = c(0, 0, 0, 0, 0, 0, 8, 0, 8, 0),
  TotalQuantity = QuantityOnHand + QuantityOrdered + OnOrder,
  OrderCost = ifelse(QuantityOrdered > 0, 10, 0), 
  # Carry cost is $1/ unit on ending balance 
  CarryCost = ifelse(QuantityOnHand > 0, QuantityOnHand * carry_cost, 0), 
  StockOutCost = ifelse(QuantityOnHand < 0, 3, 0),
  TotalCost = OrderCost + StockOutCost + CarryCost, 
  CumulativeCost = cumsum(TotalCost)
) %>% 
  knitr::kable()
```


- If the Total Quantity is less than or equal to 7, place an order. 
  - The order will have a lead time based on the lead time analysis done
  
Purpose is to balance stock outs, reorder point, and carry costs to reduce 
the cumulative total cost of managing inventory. 

**End simulation**

## Example MRP scheduling problem 

What is the projected available balance in week number 1?

```{r}
# Batch Size when placing orders 
q = 20
# Lead time 
lead_time = 1
# Safety Stock 
safety_stock = 0
# Starting balance 
starting_balance = 21 

# Work problem 
tibble(
  product = 'parta',
  week = 1:6, 
  # Given part of the problem 
  gross_requirements =        c(5, 15, 18, 8, 12, 22),
  # Starting balance (21) - gross requirements 
  ProjectedAvailableBalance = c(16, 1, 3, 15, 3, 1), 
  # Need to cover the gross requirements for the subsequent week 
  PlannedOrder =              c(0, 20, 20, 0, 20, 0),
  # One week lead time stated in problem 
  ScheduledReceived =         c(0, 0, 20, 20, 0, 20)
) %>% 
  knitr::kable()
```

If we want to reduce the number of stock outs, we can increase the quantity 
ordered. 

We calculate the gross requirements for sub-assembly b by looking at the planned 
orders for product a. 

- Sub-assemblies are derived demand from the finished product. We only need to 
build sub-assemblies when product a is ordered. 
  - Planned order in week 2, 3, and 5 will have gross requirements of: 
  
`b requirements 20 (planned order a) * 2 (each a requires two b sub-assemblies)`

```{r}
# Batch size 40 
q = 40 
# Two week lead time 
lead_time = 2 

tibble(
  product = 'partb',
  week = 1:6, 
  # Calculated by looking at the Planned order field. 
  gross_requirements =        c(0, 40, 40, 0, 40, 0),
  # Starting balance (20) + scheduled to receive 32 in first week & 0 gross req.
  ProjectedAvailableBalance = c(52, 12, 12, 12, 12, 12), 
  # Need to cover the gross requirements for the subsequent week 
  PlannedOrder =              c(40, 0, 40, 0, 0, 0),
  # Two week lead time stated in problem 
  ScheduledReceived =         c(32, 0, 40, 0, 40, 0)
) %>% 
  knitr::kable()
```



Part c: to get the gross requirements - multiply the planned order releases for 
part a and c by 1.. 


```{r}
# Lot for lot: order what is required 
q = 'LotForLot'
# One week lead time 
lead_time = 1
# Maintain safety stock of 10: Add 10 to gross_requirements 
a = c(0, 20, 20, 0, 20, 0)
b = c(40, 0, 40, 0, 0, 0)
c_gross_requirements  = a + b

tibble(
  product = 'partc',
  week = 1:6, 
  # Calculated by looking at the Planned order field. 
  gross_requirements =  c_gross_requirements, # 40, 20, 60, 0, 20, 0
  # Starting balance (50) + scheduled to receive 32 in first week & 0 gross req.
  ProjectedAvailableBalance = c(10, 10, 10, 10, 10, 10), 
  # Need to cover the gross requirements for the subsequent week 
  PlannedOrder =              c(20, 60, 0, 20, 0, 0),
  # Two week lead time stated in problem 
  ScheduledReceived =         c(0, 20, 60, 0, 20, 0)
) %>% 
  knitr::kable()
```